Automated residual plot assessment with the \textsf{R} package \textsf{autovi} and the \textsf{Shiny} application \textsf{autovi.web}

### 2025-09-06 autovi_paper.tex
Fix most paper issues

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Automated residual plot assessment with the  package  and the  app },
  pdfauthor={Weihao Li; Dianne Cook; Emi Tanaka; Susan VanderPlas; Klaus Ackermann},
  pdfkeywords={initial data analysis, statistical graphics, data
visualization, visual inference, computer vision, machine
learning, hypothesis testing, regression analysis, model diagnostics},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\title{Automated residual plot assessment with the \textsf{R} package
\textsf{autovi} and the \textsf{Shiny} app \textsf{autovi.web}}

@@

\begin{abstract}
Visual assessment of residual plots is a common approach for diagnosing
linear models, but it relies on manual evaluation, which does not scale
well and can lead to inconsistent decisions across analysts. The lineup
protocol, which embeds the observed plot among null plots, can reduce
subjectivity but requires even more human effort. In today's data-driven
world, such tasks are well-suited for automation. We present a new
\textsf{R} package that uses a computer vision model to automate the
evaluation of residual plots. An accompanying \textsf{Shiny} app is
provided for ease of use. Given a sample of residuals, the model
predicts a visual signal strength (VSS) and offers supporting
information to help analysts assess model fit.
\end{abstract}

@@

As noted in \citet{li2024automated}, the lineup protocol has significant
limitations in large-scale applications due to its reliance on human
labor. To overcome this constraint, a computer vision model was
developed alongside a corresponding statistical testing procedure to
autmoate the assessment of residual plots. The model takes as input a
residual plot and a set of auxiliary variables (such as the number of
observations) and outputs a predicted visual signal strength (VSS). This
VSS estimates the degree of deviation between the residual distribution
of the fitted model and the reference distribution expected under
correct model specification.

@@

\section{\texorpdfstring{\textsf{R} package:
autovi}{ package: autovi}}\label{sec-autovi}

The main purpose of \texttt{autovi} is to provide rejection decisions
and \(p\)-values for testing the null hypothesis (\(H_0\)) that the
regression model is correctly specified. The package provides automated
interpretation of residual plots using computer vision. The name
\texttt{autovi} stands for \textbf{auto}mated \textbf{v}isual
\textbf{i}nference. This functionality can be accessed through the
\textsf{R} package \texttt{autovi}, or through a web interface,
\texttt{autovi.web}, which enables users to perform analyses without
installing \textsf{R}, \textsf{Python}, or their associated dependencies
locally.

@@

Figure~\ref{fig-three-examples} shows three sets of plots of residuals
against fitted values. The simulated example in (a) might be interpreted
as a heteroscedastic pattern, however the automated reading would
predict this to have a visual signal strength (VSS) of 1.53, with a
corresponding \(p\)-value of 0.25. This means it would be interpreted as
a good residual plot, that there is nothing in the data to indicate a
violation of model assumptions. Skewness in the predictor variables is
generating the apparent heteroscedasticity, where the smaller variance
in residuals at larger fitted values is due to smaller sample size only.
The Breusch-Pagan test \citep{breusch1979simple} for heteroscedasticity
would also not reject this as good residual plot.

The data in (b) is generated by fitting a linear model predicting
\texttt{mpg} based on \texttt{hp} using the \texttt{datasets::mtcars}.
It is a small data set, and there is a hint of nonlinear structure not
captured by the model. The automated plot reading would predict a VSS of
3.57, which has a \(p\)-value less than 0.05. That is, the nonlinear
structure is most likely real, and indicates a problem with the model.
The conventional test, a Ramsey Regression Equation Specification Error
Test (RESET) \citep{ramsey1969tests} would also strongly detect the
nonlinearity.

The third example is generated using the \texttt{surreal} package
\citep{surreal}, where structure residuals are embedded in the data. In
this case, a quote inspired by Tukey, ``visual summaries focus on
unexpected values'', is used to define the residual structure. The
automated plot reading predicts the VSS to be 5.87, with a \(p\)-value
less than 0.05. Visually, the structure is strikingly clear, but a RESET
test for nonlinear structure would not report a problem. (It would be
detected by a Breusch-Pagan for heteroscedasticity and also Shapiro-Wilk
test \citep{shapiro1965analysis} for non-normality.)

@@

\caption{\label{fig-three-examples}Reading residual plots can be a
difficult task, particularly for students new to statistical modeling.
The \texttt{autovi} package makes it easier. Here are three examples of
residual plots, which may appear to have structure. According to autovi,
the visual signal strengths (VSS) of these three examples are
approximately (a) 1.53, (b) 3.57, (c) 5.87, resulting in (b), (c) being
significant violations of good residuals, but (a) is consistent with a
good residual plot.}

@@

The \texttt{autovi} package is built on the \texttt{bandicoot}
object-oriented programming (OOP) system \citep{bandicoot}, marking a
departure from R's traditional S3 generic system. This OOP architecture
enhances flexibility and modularity, allowing users to redefine key
functions through method overriding.

The \texttt{autovi} infrastructure effectively integrates multiple
programming languages and libraries into a comprehensive analytical
tool. It relies on five core libraries from \textsf{Python} and
\textsf{R}, each playing a critical role in the analysis pipeline. In
\textsf{Python}, \texttt{pillow} \citep{clark2015pillow} handles image
processing tasks such as reading and resizing PNG files of residual
plots, then converting them into input tensors for further analysis.
\texttt{TensorFlow} \citep{abadi2016tensorflow}, a key component of
modern machine learning, is used to predict the VSS of these plots using
a pre-trained convolutional neural network.

In the \textsf{R} environment, \texttt{autovi} utilizes several
libraries. \texttt{ggplot2} \citep{ggplot2} generates the initial
residual plots, saved as PNG files for visual input. \texttt{cassowaryr}
\citep{mason2022cassowaryr} computes scagnostics (scatter plot
diagnostics), providing numerical features that capture statistical
properties of the plots. These scagnostics complement the visual
analysis by offering quantitative metrics as secondary input to the
computer vision model. \texttt{reticulate} \citep{reticulate} enables
seamless communication between \textsf{R} and \textsf{Python}.

@@

The \texttt{autovi} package is available on CRAN. It is actively
developed and maintained, with the latest updates accessible on GitHub.
This paper uses \texttt{autovi} version 0.4.2. The package includes
internal functions to check the current \textsf{Python} environment used
by the \texttt{reticulate} package. If the necessary \textsf{Python}
packages are not installed in the \textsf{Python} interpreter, an error
will be raised. If you want to select a specific \textsf{Python}
environment, you can do so by calling the
\texttt{reticulate::use\_python()} function before using the
\texttt{autovi} package.

We recommend using the \textsf{Shiny} app \texttt{autovi.web} if users
encounter installation problems.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Load the \texttt{autovi} package using the \texttt{library()}
  function.
\item
  Create a checker object with a linear regression model.
\item
  Call the \texttt{check()} method of the checker, which, by default,
  predicts the VSS for the true residual plot, 100 null plots, and 100
  bootstrapped plots. The method stores the predictions internally and
  prints a concise results report.
\end{enumerate}

@@

\includegraphics[width=1\textwidth,height=\textheight]{autovi_paper_files/figure-pdf/fig-autovi-diag-1.png}

@@

\caption{\label{fig-autovi-diag}Diagram illustrating the infrastructure
of the extsf\{R\} package \texttt{autovi}. The modules in green are
primary inputs provided by users. Modules in blue are overridable
methods that can be modified to accommodate users' specific needs. The
module in yellow is a pre-defined non-overridable method. The modules in
red are primary outputs of the package.}

@@

The initial motivation for developing \texttt{autovi} was to create a
convenient interface for sharing the models described and trained in
\citet{li2024automated}. However, recognizing that the classical normal
linear regression model represents a restricted class of models, we
sought to avoid limiting the potential for future extensions, whether by
the original developers or other developers. As a result, the package
was designed to function seamlessly with linear regression models with
minimal modification and few required arguments, while also
accommodating other classes of models through partial infrastructure
substitution. This modular and customizable design allows
\texttt{autovi} to handle a wide range of residual diagnostics tasks.

The infrastructure of \texttt{autovi} consists of ten core modules: data
extraction, bootstrapping and model refitting, fitted values and
residuals extraction, auxiliary computation, null residual simulation,
plotting, plot saving, image reading and resizing, VSS prediction, and
\(p\)-value computation. Each module is designed with minimal dependency
on the preceding modules, allowing users to customize parts of the
infrastructure without affecting its overall integrity. An overview of
this infrastructure is illustrated in Figure~\ref{fig-autovi-diag}.

The package takes regression models and a \texttt{Keras} model as
primary inputs. Modules for VSS prediction and \(p\)-value computation
are fixed but accessible via function arguments, using
\texttt{TensorFlow} for inference and statistical testing. The image
loading module is also fixed, using \texttt{PIL} to read and resize
images based on the \texttt{Keras} model's input shape. The remaining
seven modules are overridable, allowing users to adapt the workflow as
needed. The data extraction module extracts a \texttt{data.frame}
containing variables used in the regression model. The bootstrapping and
refitting module resamples the data and refits the model. The fitted
values and residuals extraction module returns these values as a
\texttt{data.frame}. The auxiliary computation module calculates
scagnostics such as monotonicity. The plotting module generates a
\texttt{ggplot} in a standard format, and the plot saving module exports
it at the same resolution as the training images. These modules are
described in detail in the package documentation.

\subsection{Extension to Other Model
Classes}\label{extension-to-other-model-classes}

The \texttt{autovi} \textsf{R} package can be extended to accommodate
other classes of models beyond linear regression, such as generalized
linear models (\texttt{glm}). This is achieved by substituting the
relevant overridable modules, and if needed, supplying a different
\texttt{Keras} model.

@@

It is important to note, however, that the pre-trained computer vision
model included in \texttt{autovi}, such as \texttt{vss\_phn\_32} (see
\texttt{list\_keras\_models()} for the full list of available models),
was developed specifically for diagnostics of linear regression. Its
applicability to other model classes relies on the assumption that the
null residual plots exhibit characteristics broadly consistent with
those of well-behaved linear regression residuals, that is, residuals
should be approximately randomly scattered around zero, display roughly
constant variance across the range of fitted values, and exhibit no
discernible structure or curvature. If these conditions are not met, or
if model violations do not give rise to visually detectable patterns,
the validity of the automated diagnostics may be compromised. In such
cases, users are encouraged to train and apply their own \texttt{Keras}
models. Detailed guidance on model training and discussion on extending
the methodology to other model classes can be found in
\citet{li2024automated}.

\section{Web interface: autovi.web}\label{sec-autovi-web}

The \texttt{autovi.web} \textsf{Shiny} application extends the
functionality of \texttt{autovi} by offering a user-friendly web
interface for automated residual plot assessment. This eliminates the
common challenges associated with software installation, so users can
avoid managing \textsf{Python} environments or handling version
requirements for \textsf{R} libraries. The platform is cross-platform
and accessible on various devices and operating systems, making it
suitable even for users without \textsf{R} programming experience.
Additionally, updates are managed centrally, ensuring that users always
have access to the latest features. This section discusses the
implementation based on \texttt{autovi.web} version 0.1.0.

@@

The interface \texttt{autovi.web} is built using the \texttt{shiny}
\citep{shiny} and \texttt{shinydashboard} \citep{shinydashboard}
\textsf{R} packages. Hosted on the
\href{https://www.shinyapps.io}{shinyapps.io} domain, the application is
accessible through any modern web browser. The \textsf{R} packages
\texttt{htmltools} \citep{htmltools} and \texttt{shinycssloaders}
\citep{shinycssloaders} are used to render markdown documentation in
\textsf{Shiny} application, and for loading animations for
\textsf{Shiny} widgets, respectively.

Determining the best way to implement the backend was difficult. In our
initial planning for \texttt{autovi.web}, we considered implementing the
entire web application using the \texttt{webr} framework \citep{webr},
which would have allowed the entire application to run directly in the
user's browser. However, \texttt{webr} does not support packages which
use compiled fortran code, which is required by \texttt{splancs}
\citep{splancs}, a dependency of \texttt{autovi}. In the future, it is
possible that a working Emscripten \citep{zakai2011emscripten} version
of this package may allow full \texttt{webr} support.

We also explored the possibility of implementing the web interface using
frameworks built on other languages, such as \textsf{Python}. However,
server hosting domains that natively support \textsf{Python} servers
typically do not have the latest version of \textsf{R} installed.
Additionally, calling \textsf{R} from \textsf{Python} is typically done
using the \texttt{rpy2} \textsf{Python} library \citep{rpy2}, but this
approach can be awkward when dealing with language syntax related to
non-standard evaluation. Another option we considered was renting a
server where we could have full control, such as those provided by cloud
platforms like Google Cloud Platform (GCP) or Amazon Web Services (AWS).
However, deploying and maintaining the server securely requires some
expertise. Ultimately, the most practical solution was to use the
\texttt{shiny} and \texttt{shinydashboard} frameworks, which are
well-established in the \textsf{R} community and offer a solid
foundation for web application development.

The server-side configuration of \texttt{autovi.web} is carefully
designed to support its functionality. Most required \textsf{Python}
libraries, including \texttt{pillow} and \texttt{numpy}, are
pre-installed on the server. These libraries are integrated into the
\textsf{Shiny} application using the \texttt{reticulate} package, which
provides an interface between \textsf{R} and \textsf{Python}.

Due to shinyapps.io's resource policy, inactive servers enter sleep
mode, clearing the local \textsf{Python} environment. When reactivated
for a new session, libraries must be reinstalled. While this ensures a
clean environment for each session, it may lead to slightly longer
loading times for the first user after a period of inactivity.

In contrast to \texttt{autovi}, \texttt{autovi.web} leverages
\texttt{TensorFlow.js}, a JavaScript library that allows the execution
of machine learning models directly in the browser. This choice enables
native browser execution, enhancing compatibility across different user
environments, and shifts the computational load from the server to the
client-side. \texttt{TensorFlow.js} also offers better scalability and
performance, especially when dealing with resource-intensive computer
vision models on the web.

While \texttt{autovi} requires downloading the pre-trained computer
vision models from GitHub, these models in ``.keras'' file format are
incompatible with \texttt{TensorFlow.js}. Therefore, we extract and
store the model weights in JSON files and include them as extra
resources in the \textsf{Shiny} application. When the application
initializes, \texttt{TensorFlow.js} rebuilds the computer vision model
using these pre-stored weights.

To allow communication between \texttt{TensorFlow.js} and other
components of the \textsf{Shiny} application, the \texttt{shinyjs}
\textsf{R} package \citep{shinyjs} is used. This package allows calling
custom JavaScript code within the \textsf{Shiny} framework. The
specialized JavaScript code for initializing \texttt{TensorFlow.js} and
calling \texttt{TensorFlow.js} for VSS prediction is deployed alongside
the \textsf{Shiny} application as additional resources.

@@

The workflow of \texttt{autovi.web} is designed to be straightforward,
with numbered steps displayed in each panel. There are two example
datasets provided by the web application. The single residual plot
example uses the \texttt{dino} dataset from the \textsf{R} package
\texttt{datasauRus} \citep{datasaurus}. The lineup example uses
residuals from a simulated regression model that has a non-linearity
issue. We walk through the lineup example to further demonstrate the
workflow of the web application.

@@

The user can select to upload data as either a single set of residuals
and fitted values in a two (or more) column CSV file or a pre-computed
lineup of residuals and null datasets in a three (or more) column CSV
file (i.e.~multiple sets of residuals and fitted values with a column
indicating the set label). Here we illustrate use with lineup example
data sets (Figure~\ref{fig-autovi-web-setup}). To use the lineup example
data, click the ``Use Lineup Example'' button. The data status will then
update to show the number of rows and columns in the dataset, and the
CSV type will automatically be selected to the correct option. Since the
example dataset follows the variable naming conventions assumed by the
web application, the columns for fitted values, residuals, and labels of
residual plots are automatically mapped such that the column named as
\texttt{.fitted} is mapped to fitted values, \texttt{.resid} is mapped
to residuals and if applicable, \texttt{.sample} to labels of the
residual set (middle image). If the user is working with a custom
dataset, these options must be set accordingly. Whenever a data
containing a lineup, the user must manually select the label for the
true residual plot, otherwise the web application does not provide all
the results. The last step is to click the play button (right image) to
start the assessment.

@@

\includegraphics[width=1\textwidth,height=\textheight]{autovi_paper_files/figure-pdf/fig-autovi-web-setup-1.png}

@@

\caption{\label{fig-autovi-web-setup}To begin the workflow for
\texttt{autovi} using the lineup example dataset, the user clicks the
``Use Lineup Example'' button (left) to load the example dataset, during
which the data status and CSV type will be automatically updated. The
user must manually select the label for the true residual plot (middle)
to compute further results. The user initiates the assessment of the
lineup example data by clicking the run button (right).}

@@

\includegraphics[width=1\textwidth,height=\textheight]{autovi_paper_files/figure-pdf/fig-autovi-web-lineup-1.png}

@@

\includegraphics[width=1\textwidth,height=\textheight]{autovi_paper_files/figure-pdf/fig-autovi-web-distributions-1.png}

@@

This paper presents new regression diagnostics software, the \textsf{R}
package \texttt{autovi} and its accompanying web interface,
\texttt{autovi.web}. It addresses a critical gap in the current
landscape of statistical software. While regression tools are widely
available, effective and efficient diagnostic methods have lagged
behind, particularly in the field of residual plot interpretation.

The \texttt{autovi} \textsf{R} package, introduced in this paper,
automates the assessment of residual plots by incorporating a computer
vision model, reducing reliance on time-consuming and potentially
inconsistent human interpretation. This automation improves the
efficiency of the diagnostic process and promotes consistency in model
evaluation across different users and studies.

>>>>

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Automated residual plot assessment with the  package  and the  application },
  pdfauthor={Weihao Li; Dianne Cook; Emi Tanaka; Susan VanderPlas; Klaus Ackermann},
  pdfkeywords={initial data analysis, statistical graphics, data
visualization, visual inference, computer vision, machine
learning, hypothesis testing, regression analysis, model diagnostics},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\title{Automated residual plot assessment with the \textsf{R} package
\textsf{autovi} and the \textsf{Shiny} application \textsf{autovi.web}}

@@

\begin{abstract}
Visual assessment of residual plots is a common approach for diagnosing
linear models, but it relies on manual evaluation, which does not scale
well and can lead to inconsistent decisions across analysts. The lineup
protocol, which embeds the observed plot among null plots, can reduce
subjectivity but requires even more human effort. In today's data-driven
world, such tasks are well-suited for automation. We present a new
\textsf{R} package that uses a computer vision model to automate the
evaluation of residual plots. An accompanying \textsf{Shiny} application
is provided for ease of use. Given a sample of residuals, the model
predicts a visual signal strength (VSS) and offers supporting
information to help analysts assess model fit.
\end{abstract}

@@

As noted in \citet{li2024automated}, the lineup protocol has significant
limitations in large-scale applications due to its reliance on human
labor. To overcome this constraint, a computer vision model was
developed alongside a corresponding statistical testing procedure to
automate the assessment of residual plots. The model takes as input a
residual plot and a set of auxiliary variables (such as the number of
observations) and outputs a predicted visual signal strength (VSS). This
VSS estimates the degree of deviation between the residual distribution
of the fitted model and the reference distribution expected under
correct model specification.

@@

\section{\texorpdfstring{\textsf{R} package:
\textsf{autovi}}{ package: }}\label{sec-autovi}

The main purpose of \textsf{autovi} is to provide rejection decisions
and \(p\)-values for testing the null hypothesis (\(H_0\)) that the
regression model is correctly specified. The package provides automated
interpretation of residual plots using computer vision. The name
\textsf{autovi} stands for \textbf{auto}mated \textbf{v}isual
\textbf{i}nference. This functionality can be accessed through the
\textsf{R} package \textsf{autovi}, or through a web interface,
\textsf{autovi.web}, which enables users to perform analyses without
installing \textsf{R}, \textsf{Python}, or their associated dependencies
locally.

@@

Figure~\ref{fig-three-examples} shows three sets of plots of residuals
against fitted values. The simulated example in (a) might be interpreted
as a heteroscedastic pattern, however the automated reading would
predict this to have a visual signal strength (VSS) of 1.53, with a
corresponding \(p\)-value of 0.25. This means it would be interpreted as
a good residual plot, that there is nothing in the data to indicate a
violation of model assumptions. Skewness in the predictor variables is
generating the apparent heteroscedasticity, where the smaller variance
in residuals at larger fitted values is due to smaller sample size only.
The Breusch--Pagan test \citep{breusch1979simple} for heteroscedasticity
would also not reject this as good residual plot.

The data in (b) is generated by fitting a linear model predicting
\texttt{mpg} based on \texttt{hp} using the
\textsf{datasets}\texttt{::mtcars}. It is a small data set, and there is
a hint of nonlinear structure not captured by the model. The automated
plot reading would predict a VSS of 3.57, which has a \(p\)-value less
than 0.05. That is, the nonlinear structure is most likely real, and
indicates a problem with the model. The conventional test, a Ramsey
Regression Equation Specification Error Test (RESET)
\citep{ramsey1969tests} would also strongly detect the nonlinearity.

The third example is generated using the \textsf{surreal} package
\citep{surreal}, where structure residuals are embedded in the data. In
this case, a quote inspired by Tukey, `visual summaries focus on
unexpected values', is used to define the residual structure. The
automated plot reading predicts the VSS to be 5.87, with a \(p\)-value
less than 0.05. Visually, the structure is strikingly clear, but a RESET
test for nonlinear structure would not report a problem. (It would be
detected by a Breusch--Pagan for heteroscedasticity and also
Shapiro--Wilk test \citep{shapiro1965analysis} for non-normality.)

@@

\caption{\label{fig-three-examples}Reading residual plots can be a
difficult task, particularly for students new to statistical modeling.
The \textsf{autovi} package makes it easier. Here are three examples of
residual plots, which may appear to have structure. According to autovi,
the visual signal strengths (VSS) of these three examples are
approximately (a) 1.53, (b) 3.57, (c) 5.87, resulting in (b), (c) being
significant violations of good residuals, but (a) is consistent with a
good residual plot.}

@@

The \textsf{autovi} package is built on the \textsf{bandicoot}
object-oriented programming (OOP) system \citep{bandicoot}, marking a
departure from \textsf{R}'s traditional S3 generic system. This OOP
architecture enhances flexibility and modularity, allowing users to
redefine key functions through method overriding.

The \textsf{autovi} infrastructure effectively integrates multiple
programming languages and libraries into a comprehensive analytical
tool. It relies on five core libraries from \textsf{Python} and
\textsf{R}, each playing a critical role in the analysis pipeline. In
\textsf{Python}, \textsf{pillow} \citep{clark2015pillow} handles image
processing tasks such as reading and resizing PNG files of residual
plots, then converting them into input tensors for further analysis.
\textsf{TensorFlow} \citep{abadi2016tensorflow}, a key component of
modern machine learning, is used to predict the VSS of these plots using
a pre-trained convolutional neural network.

In the \textsf{R} environment, \textsf{autovi} utilizes several
libraries. \textsf{ggplot2} \citep{ggplot2} generates the initial
residual plots, saved as PNG files for visual input. \textsf{cassowaryr}
\citep{mason2022cassowaryr} computes scagnostics (scatter plot
diagnostics), providing numerical features that capture statistical
properties of the plots. These scagnostics complement the visual
analysis by offering quantitative metrics as secondary input to the
computer vision model. \textsf{reticulate} \citep{reticulate} enables
seamless communication between \textsf{R} and \textsf{Python}.

@@

The \textsf{autovi} package is available on CRAN. It is actively
developed and maintained, with the latest updates accessible on GitHub.
This paper uses \textsf{autovi} version 0.4.2. The package includes
internal functions to check the current \textsf{Python} environment used
by the \textsf{reticulate} package. If the necessary \textsf{Python}
packages are not installed in the \textsf{Python} interpreter, an error
will be raised. If you want to select a specific \textsf{Python}
environment, you can do so by calling the
\textsf{reticulate}\texttt{::use\_python()} function before using the
\textsf{autovi} package.

We recommend using the \textsf{Shiny} application \textsf{autovi.web} if
users encounter installation problems.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Load the \textsf{autovi} package using the \texttt{library()}
  function.
\item
  Create a checker object with a linear regression model.
\item
  Call the \texttt{check()} method of the checker, which, by default,
  predicts the VSS for the true residual plot, 100 null plots, and 100
  bootstrapped plots. The method stores the predictions internally and
  prints a concise results report.
\end{enumerate}

@@

\includegraphics[width=1\textwidth,height=\textheight]{figures/autovi.pdf}

@@

\caption{\label{fig-autovi-diag}Diagram illustrating the infrastructure
of the \textsf{R} package \textsf{autovi}. The modules in green are
primary inputs provided by users. Modules in blue are overridable
methods that can be modified to accommodate users' specific needs. The
module in yellow is a pre-defined non-overridable method. The modules in
red are primary outputs of the package.}

@@

The initial motivation for developing \textsf{autovi} was to create a
convenient interface for sharing the models described and trained in
\citet{li2024automated}. However, recognizing that the classical normal
linear regression model represents a restricted class of models, we
sought to avoid limiting the potential for future extensions, whether by
the original developers or other developers. As a result, the package
was designed to function seamlessly with linear regression models with
minimal modification and few required arguments, while also
accommodating other classes of models through partial infrastructure
substitution. This modular and customisable design allows
\textsf{autovi} to handle a wide range of residual diagnostics tasks.

The infrastructure of \textsf{autovi} consists of ten core modules: data
extraction, bootstrapping and model refitting, fitted values and
residuals extraction, auxiliary computation, null residual simulation,
plotting, plot saving, image reading and resizing, VSS prediction, and
\(p\)-value computation. Each module is designed with minimal dependency
on the preceding modules, allowing users to customize parts of the
infrastructure without affecting its overall integrity. An overview of
this infrastructure is illustrated in Figure~\ref{fig-autovi-diag}.

The package takes regression models and a \textsf{Keras} model as
primary inputs. Modules for VSS prediction and \(p\)-value computation
are fixed but accessible via function arguments, using
\textsf{TensorFlow} for inference and statistical testing. The image
loading module is also fixed, using \textsf{PIL} to read and resize
images based on the \textsf{Keras} model's input shape. The remaining
seven modules are overridable, allowing users to adapt the workflow as
needed. The data extraction module extracts a \texttt{data.frame}
containing variables used in the regression model. The bootstrapping and
refitting module resamples the data and refits the model. The fitted
values and residuals extraction module returns these values as a
\texttt{data.frame}. The auxiliary computation module calculates
scagnostics such as monotonicity. The plotting module generates a
\texttt{ggplot} in a standard format, and the plot saving module exports
it at the same resolution as the training images. These modules are
described in detail in the package documentation.

\subsection{Extension to other model
classes}\label{extension-to-other-model-classes}

The \textsf{autovi} \textsf{R} package can be extended to accommodate
other classes of models beyond linear regression, such as generalized
linear models (\texttt{glm}). This is achieved by substituting the
relevant overridable modules, and if needed, supplying a different
\textsf{Keras} model.

@@

It is important to note, however, that the pre-trained computer vision
model included in \textsf{autovi}, such as \texttt{vss\_phn\_32} (see
\texttt{list\_keras\_models()} for the full list of available models),
was developed specifically for diagnostics of linear regression. Its
applicability to other model classes relies on the assumption that the
null residual plots exhibit characteristics broadly consistent with
those of well-behaved linear regression residuals, that is, residuals
should be approximately randomly scattered around zero, display roughly
constant variance across the range of fitted values, and exhibit no
discernible structure or curvature. If these conditions are not met, or
if model violations do not give rise to visually detectable patterns,
the validity of the automated diagnostics may be compromised. In such
cases, users are encouraged to train and apply their own \textsf{Keras}
models. Detailed guidance on model training and discussion on extending
the methodology to other model classes can be found in
\citet{li2024automated}.

\section{\texorpdfstring{Web interface:
\textsf{autovi.web}}{Web interface: }}\label{sec-autovi-web}

The \textsf{autovi.web} \textsf{Shiny} application extends the
functionality of \textsf{autovi} by offering a user-friendly web
interface for automated residual plot assessment. This eliminates the
common challenges associated with software installation, so users can
avoid managing \textsf{Python} environments or handling version
requirements for \textsf{R} libraries. The platform is cross-platform
and accessible on various devices and operating systems, making it
suitable even for users without \textsf{R} programming experience.
Additionally, updates are managed centrally, ensuring that users always
have access to the latest features. This section discusses the
implementation based on \textsf{autovi.web} version 0.1.0.

@@

The interface \textsf{autovi.web} is built using the \textsf{shiny}
\citep{shiny} and \textsf{shinydashboard} \citep{shinydashboard}
\textsf{R} packages. Hosted on the
\href{https://www.shinyapps.io}{shinyapps.io} domain, the application is
accessible through any modern web browser. The \textsf{R} packages
\textsf{htmltools} \citep{htmltools} and \textsf{shinycssloaders}
\citep{shinycssloaders} are used to render markdown documentation in
\textsf{Shiny} application, and for loading animations for
\textsf{Shiny} widgets, respectively.

Determining the best way to implement the backend was difficult. In our
initial planning for \textsf{autovi.web}, we considered implementing the
entire web application using the \textsf{webr} framework \citep{webr},
which would have allowed the entire application to run directly in the
user's browser. However, \textsf{webr} does not support packages which
use compiled \textsf{Fortran} code, which is required by
\textsf{splancs} \citep{splancs}, a dependency of \textsf{autovi}. In
the future, it is possible that a working \textsf{Emscripten}
\citep{zakai2011emscripten} version of this package may allow full
\textsf{webr} support.

We also explored the possibility of implementing the web interface using
frameworks built on other languages, such as \textsf{Python}. However,
server hosting domains that natively support \textsf{Python} servers
typically do not have the latest version of \textsf{R} installed.
Additionally, calling \textsf{R} from \textsf{Python} is typically done
using the \textsf{rpy2} \textsf{Python} library \citep{rpy2}, but this
approach can be awkward when dealing with language syntax related to
non-standard evaluation. Another option we considered was renting a
server where we could have full control, such as those provided by cloud
platforms like Google Cloud Platform (GCP) or Amazon Web Services (AWS).
However, deploying and maintaining the server securely requires some
expertise. Ultimately, the most practical solution was to use the
\textsf{shiny} and \textsf{shinydashboard} frameworks, which are
well-established in the \textsf{R} community and offer a solid
foundation for web application development.

The server-side configuration of \textsf{autovi.web} is carefully
designed to support its functionality. Most required \textsf{Python}
libraries, including \textsf{pillow} and \textsf{numpy}, are
pre-installed on the server. These libraries are integrated into the
\textsf{Shiny} application using the \textsf{reticulate} package, which
provides an interface between \textsf{R} and \textsf{Python}.

Due to \href{https://www.shinyapps.io}{shinyapps.io}'s resource policy,
inactive servers enter sleep mode, clearing the local \textsf{Python}
environment. When reactivated for a new session, libraries must be
reinstalled. While this ensures a clean environment for each session, it
may lead to slightly longer loading times for the first user after a
period of inactivity.

In contrast to \textsf{autovi}, \textsf{autovi.web} leverages
\textsf{TensorFlow.js}, a \textsf{JavaScript} library that allows the
execution of machine learning models directly in the browser. This
choice enables native browser execution, enhancing compatibility across
different user environments, and shifts the computational load from the
server to the client-side. \textsf{TensorFlow.js} also offers better
scalability and performance, especially when dealing with
resource-intensive computer vision models on the web.

While \textsf{autovi} requires downloading the pre-trained computer
vision models from GitHub, these models in `.keras' file format are
incompatible with \textsf{TensorFlow.js}. Therefore, we extract and
store the model weights in JSON files and include them as extra
resources in the \textsf{Shiny} application. When the application
initializes, \textsf{TensorFlow.js} rebuilds the computer vision model
using these pre-stored weights.

To allow communication between \textsf{TensorFlow.js} and other
components of the \textsf{Shiny} application, the \textsf{shinyjs}
\textsf{R} package \citep{shinyjs} is used. This package allows calling
custom \textsf{JavaScript} code within the \textsf{Shiny} framework. The
specialized \textsf{JavaScript} code for initializing
\textsf{TensorFlow.js} and calling \textsf{TensorFlow.js} for VSS
prediction is deployed alongside the \textsf{Shiny} application as
additional resources.

@@

The workflow of \textsf{autovi.web} is designed to be straightforward,
with numbered steps displayed in each panel. There are two example
datasets provided by the web application. The single residual plot
example uses the \texttt{dino} dataset from the \textsf{R} package
\textsf{datasauRus} \citep{datasaurus}. The lineup example uses
residuals from a simulated regression model that has a non-linearity
issue. We walk through the lineup example to further demonstrate the
workflow of the web application.

@@

The user can select to upload data as either a single set of residuals
and fitted values in a two (or more) column CSV file or a pre-computed
lineup of residuals and null datasets in a three (or more) column CSV
file (i.e.~multiple sets of residuals and fitted values with a column
indicating the set label). Here we illustrate use with lineup example
data sets (Figure~\ref{fig-autovi-web-setup}). To use the lineup example
data, click the `Use Lineup Example' button. The data status will then
update to show the number of rows and columns in the dataset, and the
CSV type will automatically be selected to the correct option. Since the
example dataset follows the variable naming conventions assumed by the
web application, the columns for fitted values, residuals, and labels of
residual plots are automatically mapped such that the column named as
\texttt{.fitted} is mapped to fitted values, \texttt{.resid} is mapped
to residuals and if applicable, \texttt{.sample} to labels of the
residual set (middle image). If the user is working with a custom
dataset, these options must be set accordingly. Whenever a data
containing a lineup, the user must manually select the label for the
true residual plot, otherwise the web application does not provide all
the results. The last step is to click the play button (right image) to
start the assessment.

@@

\includegraphics[width=1\textwidth,height=\textheight]{figures/autovi_web_setup.pdf}

@@

\caption{\label{fig-autovi-web-setup}To begin the workflow for
\textsf{autovi} using the lineup example dataset, the user clicks the
`Use Lineup Example' button (left) to load the example dataset, during
which the data status and CSV type will be automatically updated. The
user must manually select the label for the true residual plot (middle)
to compute further results. The user initiates the assessment of the
lineup example data by clicking the run button (right).}

@@

\includegraphics[width=1\textwidth,height=\textheight]{figures/autovi_web_lineup.pdf}

@@

\includegraphics[width=1\textwidth,height=\textheight]{figures/autovi_web_distributions.pdf}

@@

This paper presents new regression diagnostics software, the \textsf{R}
package \textsf{autovi} and its accompanying web interface,
\textsf{autovi.web}. It addresses a critical gap in the current
landscape of statistical software. While regression tools are widely
available, effective and efficient diagnostic methods have lagged
behind, particularly in the field of residual plot interpretation.

The \textsf{autovi} \textsf{R} package, introduced in this paper,
automates the assessment of residual plots by incorporating a computer
vision model, reducing reliance on time-consuming and potentially
inconsistent human interpretation. This automation improves the
efficiency of the diagnostic process and promotes consistency in model
evaluation across different users and studies.

The development of the accompanying \textsf{Shiny} app,
\textsf{autovi.web}, expands access to these advanced diagnostic tools,
by providing a user-friendly interface. It makes automated residual plot
assessment accessible to a broader audience, including those who may not
have extensive programming experience. This web-based solution
effectively addresses the potential barriers to adoption, such as
complex dependencies and installation requirements, that are often
associated with advanced statistical software.
